### 2015-06-02 tables/tabExtra.tex
added tables and figures used in latex files

F1 & Facebook  & friendship  & no  & the friendship network of Renato Fabbri (author)  & 1367  & 28606 \\\hline
F2 & Facebook  & friendship  & no  & the friendship network of Massimo Canevacci (senior anthropologist)  & 4764  & 59995 \\\hline
F3 & Facebook  & friendship  & no  & the friendship network of a brazilian direct democracy group  & 3599  & 59471 \\\hline
F4 & Facebook  & friendship  & no  & the friendship network of the Silicon Valley Global Network group  & 2026  & 15586 \\\hline
F5 & Participa.br  & friendship  & no  & the friendship network of a brazilian federal social participation portal  & 443  & 910 \\\hline
I1 & Facebook  & interaction  & yes  & the interaction network of the Silicon Valley Global Network group  & 104  & 154 \\\hline
I2 & Facebook  & interaction  & yes  & the interaction network of a Solidarity Economy group  & 64  & 120 \\\hline
I3 & Facebook  & interaction  & yes  & the interaction network of a brazilian direct democracy group  & 214  & 310 \\\hline
I4 & Facebook  & interaction  & yes  & the interaction network of the 'Cience with Frontiers' group  & 530  & 1658 \\\hline
I5 & Participa.br  & interaction  & yes  & the interaction network of a brazilian federal social participation portal  & 222  & 300 \\\hline
TT1 & Twitter  & retweet  & yes  & the retweet network of $\approx 22k$ tweets with the hashtag \#arenaNETmundial  & 2772  & 7222 \\\hline
TT2 & Twitter  & retweet  & yes  & same as TT1, but disconnected agents are not discarded  & 2975  & 7222 \\\hline